Distance/time to travel calculations for hospitals in Scotland

### 2021-08-04 distance_to_nearest_hospital.Rmd
reorganise the file and wrote up an introduction

library(tidyverse)  # for tidy workflow
library(opendatascot)  # importing data from ScotGov open data website
library(phsmethods)  # methods for working with PHS data
library(readxl)  # for reading excel files
library(curl)  # for downloading stuff
library(lubridate)  # dealing with dates
library(janitor)  # for cleaning column names
library(ISOweek)  # for computing date from ISO week number + year
library(sf)  # for mapping
library(ggrepel)  # for 'mapping'repelling' labels/texst in ggplot
library(patchwork)  # for assembling plots
library(extrafont)  # for working with fonts
library(openxlsx)  # for creating xlsx files
library(knitr)  # for displaying tables in rmd
library(tictoc)  # for measuring time taken for calculations
```

@@

@@

if (!file.exists("./distance_to_hospital/current-hospital_flagged20210506.csv")) download.file(url = url_hospitals, destfile = "./distance_to_hospital/current-hospital_flagged20210506.csv")
hospitals <- read_csv(file = "./distance_to_hospital/current-hospital_flagged20210506.csv")

@@

# Compute straight-line distance for each datazone to nearest hospital

For this purpose, I've removed some specialist units that operate within a larger hospital department and are listed at the same geographical grid reference! Otherwise we get two units that are equidistant for some data zones.

>>>>

library(tidyverse)  # for tidy workflow
library(readxl)  # for reading excel files
library(tictoc)  # for measuring time taken for calculations
```

@@

For investigating increased numbers of deaths at home & other non-institutional settings in Scotland, we'll take into account the distance/travel time to the nearest hospital. For this purpose, we need to compute the distance to nearest hospital.

There are two basic approaches: straight-line distance, and road distance / road travel time:

## Straight-line distance

This can be computed using OS National Grid Reference values (Easting & Northing), using the locations of hospitals, and either:

* data zones
* postcodes

Postcodes are more precise, but also more privacy-invasive when working with administrative records. For this reason, we would ideally provide the computed distances at data linkage rather than after data delivery.

Straight-line distances don't necessarily represent a realistic measure of distance or travel time to hospital, however! The main benefit is its ease of calculation.

## Road distance / travel time

This is more realistic & representative than straight-line distance, but more difficult to obtain, since road distances and road spee (used in calculating travel time) aren't openly available like the OS grid reference data are.

@@

if (!file.exists("./distance_to_hospital/current-hospital_flagged20210506.csv")) download.file(url = url_hospitals, destfile = "./distance_to_hospital/current-hospital_flagged20210506.csv")

hospitals <- read_csv(file = "./distance_to_hospital/current-hospital_flagged20210506.csv")

@@

# DataZone straight-line distance to nearest hospital

For this purpose, I've removed some duplicate entries in the list of hospitals - I think they are specialist units that operate within a larger hospital and are listed at the same geographical grid reference! Otherwise we get two units that are equidistant for some data zones.
